---
title: "vingette"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Function: eq_clean_data
Description: Cleans raw data by creating a date column with year, month and day in one date, converting
    LATITUDE and LONGITUDE columns to numeric and cleaning the LOCATION_NAME column with the eq_location_clean() function.
    Negative years have been removed.

Example:
```{r}
df <- load_data() \%>\% dplyr::filter(YEAR >= 0) \%>\%
       dplyr::mutate(date = get_date(DAY, MONTH, YEAR))
  print(class(df))
  head(df)
```

Function: eq_create_label
Description: Input a dataframe (df) and the function will output a vector with HTML column names
    Required columns: CLEAN_LOCATION_NAME, EQ_PRIMARY and DEATHS

Example:
```{r}
map2 <-load_data() \%>\%
  eq_clean_data() \%>\%
  dplyr::filter(COUNTRY == "GREECE" & lubridate::year(date) >= 2002) \%>\%
  dplyr::mutate(popup_text = eq_create_label(.)) \%>\%
  eq_map(annot_col = "popup_text")
print(map2)
```

Function: eq_location_clean
Description: This function cleans the LOCATION_NAME column by removing country name and
    converting names to title case.

Example:
```{r}
df <- load_data() \%>\% eq_location_clean()
  print(class(df))
  head(df)
```

Function: eq_map
Description: Subset of earthquake data plotted on a map using leaflet.
     Input dataframe requires LONGITUDE, LATITUDE and EQ_PRIMARY columns.
     Other required argument: annotation column ('annot_col')

Example:
```{r}
 map1 <-load_data() \%>\%
   eq_clean_data() \%>\%
   dplyr::filter(COUNTRY == "GREECE" & lubridate::year(date) >= 2002) \%>\%
   eq_map(annot_col = "date")
 print(map1)
```

Function: geom_timeline
Description: This geom charts a timeline of earthquakes for a given country
     with points along the timeline for each earthquake. The size of the point inicates
     the magnitude of the earthquake, while the color indicates the number of deaths.
     The date (x) is a required aesthetic. The country (y) is optional.

Example:
```{r}
 df <- df_earthquakes \%>\% filter(COUNTRY \%in\% c("GREECE", "USA"), YEAR > 2002)
 ggplot(df, aes(x = date, y = COUNTRY,
                color = as.numeric(TOTAL_DEATHS),
                size = as.numeric(EQ_PRIMARY),
                label = CLEAN_LOCATION_NAME)) +
   geom_timeline() +
   labs(size = "Richter scale value", color = "# deaths") +
   ggplot2::theme(panel.background = ggplot2::element_blank(),
         legend.position = "bottom",
         axis.title.y = ggplot2::element_blank()) +
   ggplot2::xlab("DATE")
```

Function: geom_timeline_label
Description: Run geom_timeline first to create timeline.
     This geom will add a vertical line and label to those points.

Example:
```{r}
ggplot(df, aes(x = date, y = COUNTRY,
               color = as.numeric(TOTAL_DEATHS),
               size = as.numeric(EQ_PRIMARY),
               label = CLEAN_LOCATION_NAME)) +
  geom_timeline() +
  labs(size = "Richter scale value", color = "# deaths") +
  ggplot2::theme(panel.background = ggplot2::element_blank(),
        legend.position = "bottom",
        axis.title.y = ggplot2::element_blank()) + ggplot2::xlab("DATE") +
  geom_timeline_label(data=df)
```

Function: GeomTimeline
Description: The GeomTimeline charts a timeline of earthquakes for a given country
     with points along the timeline for each earthquake. The size of the point inicates
     the magnitude of the earthquake, while the color indicates the number of deaths.
     The date (x) is a required aesthetic. The country (y) is optional.

Example:
```{r}
df <- df_earthquakes \%>\% filter(COUNTRY \%in\% c("GREECE", "USA"), YEAR > 2002)
 ggplot(df, aes(x = date, y = COUNTRY,
                color = as.numeric(TOTAL_DEATHS),
                size = as.numeric(EQ_PRIMARY),
                label = CLEAN_LOCATION_NAME)) +
   geom_timeline() +
   labs(size = "Richter scale value", color = "# deaths") +
   ggplot2::theme(panel.background = ggplot2::element_blank(),
         legend.position = "bottom",
         axis.title.y = ggplot2::element_blank()) +
   ggplot2::xlab("DATE")
```

Function: GeomTimelineLabel
Description: The GeomTimelineLabel will add a vertical line and label to those points.

Example:
```{r}
ggplot(df, aes(x = date, y = COUNTRY,
               color = as.numeric(TOTAL_DEATHS),
               size = as.numeric(EQ_PRIMARY),
               label = CLEAN_LOCATION_NAME)) +
  geom_timeline() +
  labs(size = "Richter scale value", color = "# deaths") +
  ggplot2::theme(panel.background = ggplot2::element_blank(),
        legend.position = "bottom",
        axis.title.y = ggplot2::element_blank()) + ggplot2::xlab("DATE") +
  geom_timeline_label(data=df)
```

Function: get_date
Description: Takes the three date-related columns and turns them into one cohesive date column. Some instances of
     DAY and MONTH are NA. NA days will be replaced with 1, NA months will be replaced with Jan.
     Negative years have been removed.

Example:
```{r}
df <- load_data() \%>\% dplyr::filter(YEAR >= 0) \%>\%
       dplyr::mutate(date = get_date(DAY, MONTH, YEAR))
  print(class(df))
  head(df)
```

Function: load_data
Description: Function loads raw, tab delimited data (from NOAA, in data folder).

Example:
```{r}
df <- load_data()
  print(class(df))
  head(df)
```

Function: plot_earthquakes_timeline
Description: This plots and formats the earthquake data.
     The plot will be saved as earthquakes_timeline.png if save_png = TRUE.

Example:
```{r}
df <- df_earthquakes \%>\%
  filter(COUNTRY \%in\% c("GREECE", "USA"), YEAR > 2002)
plot_earthquakes_timeline(df, save_png=TRUE)
```

Function: plot_earthquakes_timeline_label
Description: This plots and formats the earthquake data, with the addition of labels for the points along the timeline.
     The plot will be saved as earthquakes_timeline_label.png if save_png = TRUE.

Example:
```{r}
df <- df_earthquakes \%>\%
  filter(COUNTRY \%in\% c("GREECE", "USA"), YEAR > 2002)
plot_earthquakes_timeline_label(df, save_png=TRUE)
```
